**The Heart Lies: The Role of Interoception and Theory of Mind in Deception**

### 2022-10-04 index.Rmd
added interpretations
update

---
title: '**The Role of Interoception and Theory of Mind in Deception**'
subtitle: A Subtitle
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: no
    df_print: default
    code_folding: hide
    code_download: yes
  word_document:
    reference_docx: utils/Template_Word.docx
    highlight: pygments
    toc: no
    toc_depth: 3
    df_print: kable
    number_sections: yes
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '2'
    latex_engine: xelatex
editor_options:
  chunk_output_type: console
bibliography: utils/bibliography.bib
csl: utils/apa.csl
---

@@

```{r setup, warning=FALSE, message=TRUE, include=FALSE}
# Set up the environment (or use local alternative `source("utils/config.R")`)
source("https://raw.githubusercontent.com/RealityBending/TemplateResults/main/utils/config.R")
options(dplyr.summarise.inform = FALSE) 
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
theme_set(see::theme_modern())
```

# Introduction

```{r badges, echo=FALSE, message=TRUE, warning=FALSE, results='asis'}
# This chunk is a bit complex so don't worry about it: it's made to add badges to the HTML versions
# NOTE: You have to replace the links accordingly to have working "buttons" on the HTML versions
if (!knitr::is_latex_output() && knitr::is_html_output()) {
  cat("![Build](https://github.com/RealityBending/TemplateResults/workflows/Build/badge.svg)
      [![Website](https://img.shields.io/badge/repo-Readme-2196F3)](https://github.com/RealityBending/TemplateResults)
      [![Website](https://img.shields.io/badge/visit-website-E91E63)](https://realitybending.github.io/TemplateResults/)
      [![Website](https://img.shields.io/badge/download-.docx-FF5722)](https://github.com/RealityBending/TemplateResults/raw/main/word_and_pdf/SupplementaryMaterials.docx)
      [![Website](https://img.shields.io/badge/see-.pdf-FF9800)](https://github.com/RealityBending/TemplateResults/blob/main/word_and_pdf/SupplementaryMaterials.pdf)")
}
```

<!-- Title ideas:-->
<!-- Theory of Body and Others -->

# Packages & Data

## Packages

This document was prepared on `r format(Sys.Date())`.

```{r warning=FALSE, message=TRUE, results='asis'}
library(tidyverse)
library(patchwork)
library(glmmTMB)
library(report)
library(parameters)
library(correlation)
library(modelbased)
library(performance)
library(see)

summary(report::report(sessionInfo()))
```

## Data

@@

@@

@@

@@

@@

@@

# Full Code

The full script of executive code contained in this document is reproduced here.

```{r full_code, ref.label=knitr::all_labels(), eval=FALSE}
```

# Package References

>>>>

---
title: '**The Role of Interoception and Theory of Mind in Deception**'
author: "Dominique Makowski et al."
date: "`r Sys.Date()`"
output:
  html_document:
    theme: paper
    toc: yes
    toc_depth: 3
    df_print: default
    code_folding: hide
    code_download: yes
editor_options:
  chunk_output_type: console
---

@@

```{r setup, warning=FALSE, message=TRUE, include=FALSE}
# Set up the environment (or use local alternative `source("utils/config.R")`)
source("https://raw.githubusercontent.com/RealityBending/TemplateResults/main/utils/config.R")

options(
  digits = 3,
  mc.cores = 4,
  brms.algorithm = "sampling",
  brms.backend = "cmdstanr",
  dplyr.summarise.inform = FALSE
)

knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
theme_set(see::theme_modern())
```

```{r warning=FALSE, message=TRUE, results='asis'}
library(tidyverse)
library(patchwork)
library(glmmTMB)
library(brms)
library(easystats)

show_parameters <- function(model) {
  p <- parameters::parameters(model, effects = "fixed", component="conditional", test=c("pd")) |>
    data_relocate("pd", after=-1)
  display(p, stars=TRUE)
}
```

```{r warning=FALSE, message=TRUE, results='asis'}
summary(report::report(sessionInfo()))
```

@@

# Outliers
df$HeartRate[df$Participant == "S30"] <- NA # Extreme values
# df$Confidence[df$Participant %in% c("S9", "S29")] <- NA # Extreme responses
df$RT[df$Participant == "S13"] <- NA # Slower than the others

# Remove outlier trials
df <- df |>
  group_by(Participant) |>
  mutate(Outliers_RT = as.logical(performance::check_outliers(RT, method = "zscore", threshold = qnorm(0.99999))),
         Outliers_Physio = as.logical(performance::check_outliers(HeartRate, method = "zscore", threshold = qnorm(0.99999)))) |>
  ungroup()

@@

Percentage of Confidence data removed: `r sum(is.na(df$Confidence)) / nrow(df) * 100` %  
Percentage of RT data removed: `r sum(is.na(df$RT)) / nrow(df) * 100` %  
Percentage of Heart Rate data removed: `r sum(is.na(df$HeartRate)) / nrow(df) * 100` %

@@

# Dimension Reduction

## Theory of Mind

```{r message=FALSE, warning=FALSE}
dfsub <- df |> 
  select(Participant, 
         starts_with("YONI_"), 
         starts_with("BES_")) |> 
  select(-ends_with("Total")) |> 
  group_by(Participant) |> 
  summarise_all(mean) |> 
  select(-Participant)

parameters::n_factors(dfsub)
efa <- parameters::factor_analysis(dfsub, n=1, sort=TRUE, rotation = "oblimin")
efa

df <- cbind(df, predict(efa, newdata=df, names="ToM"))
```

## Interoception

```{r message=FALSE, warning=FALSE}
dfsub <- df |> 
  select(Participant, 
         starts_with("HCT_"), 
         starts_with("MAIA_")) |> 
  select(-ends_with("Total")) |> 
  group_by(Participant) |> 
  summarise_all(mean) |> 
  select(-Participant)

parameters::n_components(dfsub)
efa <- parameters::factor_analysis(dfsub, n=4, sort=TRUE, rotation = "oblimin")
efa
df <- cbind(df, predict(efa, newdata=df, names=c("Intero_Meta", "Intero_Listening", "Intero_Focus", "Intero_Regulation")))
```

@@

## Theory of Mind Score {.tabset}

### Confidence

```{r message=FALSE, warning=FALSE}
results <- list()
for (var in c("ToM")) {
  model <- brms::brm(as.formula(paste0("Confidence ~ Answer / (Condition / ",
                                     var,
                                     ") + (1|Participant) + (1|Item)")),
                   data = df, family = "beta", refresh=0, seed=3, iter=4000,
                   prior = set_prior("student_t(1, 0, 1)", class = "b"))

  results[[var]] <- parameters::parameters(model, effects = "fixed", component = "conditional", test = c("pd"), keep = var) |>
      as.data.frame() |> 
      separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
      mutate(
        Answer = str_remove(Answer, "b_Answer"),
        Condition = str_remove(Condition, "Condition")
      )
}

display(format_table(results$ToM, stars = TRUE))
```

### RT

```{r message=FALSE, warning=FALSE}
results <- list()
for (var in c("ToM")) {
  model <- brms::brm(as.formula(paste0("RT ~ Answer / (Condition / ",
                                     var,
                                     ") + (1|Participant) + (1|Item)")),
                   data = df, refresh=0, seed=3, iter=4000,
                   prior = set_prior("student_t(1, 0, 3)", class = "b"))

  results[[var]] <- parameters::parameters(model, effects = "fixed", component = "conditional", test = c("pd"), keep = var) |>
      as.data.frame() |> 
      separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
      mutate(
        Answer = str_remove(Answer, "b_Answer"),
        Condition = str_remove(Condition, "Condition")
      )
}

display(format_table(results$ToM, stars = TRUE))
```

### Heart Rate

```{r message=FALSE, warning=FALSE}
results <- list()
for (var in c("ToM")) {
  model <- brms::brm(as.formula(paste0("HeartRate ~ Answer / (Condition / ",
                                     var,
                                     ") + (1|Participant) + (1|Item)")),
                   data = df, refresh=0, seed=3, iter=4000,
                   prior = set_prior("student_t(1, 0, 8)", class = "b"))

  results[[var]] <- parameters::parameters(model, effects = "fixed", component = "conditional", test = c("pd"), keep = var) |>
      as.data.frame() |> 
      separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
      mutate(
        Answer = str_remove(Answer, "b_Answer"),
        Condition = str_remove(Condition, "Condition")
      )
}

display(format_table(results$ToM, stars = TRUE))
```

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  select(Participant,
         starts_with("LIE_"),
         starts_with("ToM")) |>
  group_by(Participant) |>
  summarise_all(mean)

correlation(select(dfsub, starts_with("LIE_")), select(dfsub, starts_with("ToM")), bayesian=TRUE)
```

@@

## Interoception Scores {.tabset}

### Confidence

```{r message=FALSE, warning=FALSE}
results <- list()
for (var in c("Intero_Meta", "Intero_Listening", "Intero_Focus", "Intero_Regulation")) {
  model <- brms::brm(as.formula(paste0(
    "Confidence ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df, family = "beta", refresh = 0, seed=3, iter = 4000,
  prior = set_prior("student_t(1, 0, 1)", class = "b")
  )

  results[[var]] <- parameters::parameters(model, effects = "fixed", component = "conditional", test = c("pd"), keep = var) |>
      as.data.frame() |> 
      separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
      mutate(
        Answer = str_remove(Answer, "b_Answer"),
        Condition = str_remove(Condition, "Condition")
      )
}

display(format_table(results$Intero_Meta, stars = TRUE))
display(format_table(results$Intero_Listening, stars = TRUE))
display(format_table(results$Intero_Focus, stars = TRUE))
display(format_table(results$Intero_Regulation, stars = TRUE))
```

### RT

```{r message=FALSE, warning=FALSE}
results <- list()
for (var in c("Intero_Meta", "Intero_Listening", "Intero_Focus", "Intero_Regulation")) {
  model <- brms::brm(as.formula(paste0(
    "RT ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df, refresh = 0, seed=3, iter = 4000,
  prior = set_prior("student_t(1, 0, 3)", class = "b")
  )

  results[[var]] <- parameters::parameters(model, effects = "fixed", component = "conditional", test = c("pd"), keep = var) |>
      as.data.frame() |> 
      separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
      mutate(
        Answer = str_remove(Answer, "b_Answer"),
        Condition = str_remove(Condition, "Condition")
      )
}

display(format_table(results$Intero_Meta, stars = TRUE))
display(format_table(results$Intero_Listening, stars = TRUE))
display(format_table(results$Intero_Focus, stars = TRUE))
display(format_table(results$Intero_Regulation, stars = TRUE))
```

### Heart Rate

```{r message=FALSE, warning=FALSE}
results <- list()
for (var in c("Intero_Meta", "Intero_Listening", "Intero_Focus", "Intero_Regulation")) {
  model <- brms::brm(as.formula(paste0(
    "HeartRate ~ Answer / (Condition / ",
    var,
    ") + (1|Participant) + (1|Item)"
  )),
  data = df, refresh = 0, seed=3, iter = 4000,
  prior = set_prior("student_t(1, 0, 8)", class = "b")
  )

  results[[var]] <- parameters::parameters(model, effects = "fixed", component = "conditional", test = c("pd"), keep = var) |>
      as.data.frame() |> 
      separate(Parameter, sep = ":", into = c("Answer", "Condition", "Variable")) |>
      mutate(
        Answer = str_remove(Answer, "b_Answer"),
        Condition = str_remove(Condition, "Condition")
      )
}

display(format_table(results$Intero_Meta, stars = TRUE))
display(format_table(results$Intero_Listening, stars = TRUE))
display(format_table(results$Intero_Focus, stars = TRUE))
display(format_table(results$Intero_Regulation, stars = TRUE))
```

### Correlation with LIE Scale

```{r message=FALSE, warning=FALSE}
dfsub <- df |>
  select(
    Participant,
    starts_with("LIE_"),
    starts_with("Intero_")
  ) |>
  group_by(Participant) |>
  summarise_all(mean)

correlation(select(dfsub, starts_with("LIE_")), select(dfsub, starts_with("Intero_")), bayesian = TRUE)
```

@@

# Figure

```{r warning=FALSE, message=FALSE, eval=FALSE}
data <- df |>
  group_by(Participant, Answer, Condition) |> 
  select(Confidence, RT, HeartRate, ToM, Intero_Regulation, Intero_Focus, Intero_Meta, Intero_Listening) |> 
  summarise_all(.funs = list(Mean = function(x) mean(x, na.rm = TRUE), 
                             low = function(x) (mean(x, na.rm=TRUE) - sd(x, na.rm=TRUE) / 2),
                             high = function(x) (mean(x, na.rm=TRUE) + sd(x, na.rm=TRUE) / 2))) |> 
  mutate(ToM_Mean = ifelse(Condition == "Polygraph", ToM_Mean-0.01, ToM_Mean+0.01),
         Intero_Regulation_Mean = ifelse(Condition == "Polygraph", Intero_Regulation_Mean-0.01, Intero_Regulation_Mean+0.01),
         Intero_Focus_Mean = ifelse(Condition == "Polygraph", Intero_Focus_Mean-0.01, Intero_Focus_Mean+0.01),
         Intero_Meta_Mean = ifelse(Condition == "Polygraph", Intero_Meta_Mean-0.01, Intero_Meta_Mean+0.01),
         Intero_Listening_Mean = ifelse(Condition == "Polygraph", Intero_Listening_Mean-0.01, Intero_Listening_Mean+0.01))

model <- glmmTMB(Confidence ~ Answer / Condition / ToM + (1|Participant) + (1|Item), 
                 data = df, 
                 family = beta_family())

model <- brms::brm(Confidence ~ Answer / Condition / ToM + (1|Participant) + (1|Item), 
                   data = df, family = "beta", seed=3, refresh=0,
                   prior = set_prior("student_t(1, 0, 1)", class = "b"))

link_data <- estimate_relation(model, at = c("Condition", "Answer", "ToM"), length = 30)

p1 <- ggplot(link_data, aes(x = ToM, y = Predicted)) +
  geom_segment(data = data,
                  aes(x = ToM_Mean,
                      xend = ToM_Mean,
                      y = Confidence_low,
                      yend = Confidence_high,
                      color = Condition),
                  alpha=1/3) +
  geom_point2(data = data,
                  aes(x = ToM_Mean,
                      y = Confidence_Mean,
                      color = Condition),
              size=2) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Condition), alpha = 1/3) +
  geom_line(aes(color = Condition), size=1) +
  geom_text(data=data.frame(ToM = -0.3, Predicted = 0.45, Condition = "Polygraph", Answer = "Lie"),
            label = "***",
            color = "#FF5722",
            size = 10, show.legend = FALSE, fontface = "bold") +
  labs(y = "Confidence",
       x = "Theory of Mind") +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  scale_fill_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  facet_wrap(~Answer) +
  theme_modern(axis.title.space = 10) +
  theme(strip.text = element_text(size = 10),
        strip.background = element_rect(fill = c("grey"), color = "white"))
p1

model <- brms::brm(Confidence ~ Answer / Condition / Intero_Meta + (1|Participant) + (1|Item), 
                   data = df, family = "beta", seed=3, refresh=0,
                   prior = set_prior("student_t(1, 0, 1)", class = "b"))

link_data <- estimate_relation(model, at = c("Condition", "Answer", "Intero_Meta"), length = 30)

p2 <- ggplot(link_data, aes(x = Intero_Meta, y = Predicted)) +
  geom_segment(data = data,
                  aes(x = Intero_Meta_Mean,
                      xend = Intero_Meta_Mean,
                      y = Confidence_low,
                      yend = Confidence_high,
                      color = Condition),
                  alpha=1/3) +
  geom_point2(data = data,
                  aes(x = Intero_Meta_Mean,
                      y = Confidence_Mean,
                      color = Condition),
              size=2) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Condition), alpha = 1/3) +
  geom_line(aes(color = Condition), size=1) +
  geom_text(data=data.frame(Intero_Meta = -0.15, Predicted = 0.25, Condition = "Polygraph", Answer = "Lie"),
            label = "***",
            color = "#FF5722",
            size = 10, show.legend = FALSE, fontface = "bold") +
  labs(y = "Confidence",
       x = expression(Interoception["  (.97 MAIA Attention regulation, .63 MAIA Self-regulation, .60 MAIA Emotional Awareness)"])) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  scale_fill_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  facet_wrap(~Answer) +
  theme_modern(axis.title.space = 10) +
  theme(strip.text = element_text(size = 10),
        strip.background = element_rect(fill = c("grey"), color = "white"))
p2

model <- brms::brm(Confidence ~ Answer / Condition / Intero_Listening + (1|Participant) + (1|Item), 
                   data = df, family = "beta", seed=3, refresh=0,
                   prior = set_prior("student_t(1, 0, 1)", class = "b"))
link_data <- estimate_relation(model, at = c("Condition", "Answer", "Intero_Listening"), length = 30)

p3 <- ggplot(link_data, aes(x = Intero_Listening, y = Predicted)) +
  geom_segment(data = data,
                  aes(x = Intero_Listening_Mean,
                      xend = Intero_Listening_Mean,
                      y = Confidence_low,
                      yend = Confidence_high,
                      color = Condition),
                  alpha=1/3) +
  geom_point2(data = data,
                  aes(x = Intero_Listening_Mean,
                      y = Confidence_Mean,
                      color = Condition),
              size=2) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Condition), alpha = 1/3) +
  geom_line(aes(color = Condition), size=1) +
  geom_text(data=data.frame(Intero_Listening = 0.2, Predicted = 0.45, Condition = "Polygraph", Answer = "Lie"),
            label = "***",
            color = "#FF5722",
            size = 10, show.legend = FALSE, fontface = "bold") +
  labs(y = "Confidence",
       x = expression(Interoception["  (.92 MAIA Body listening, -.60 HCT Awareness, .53 HCT Trusting)"])) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  scale_fill_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  facet_wrap(~Answer) +
  theme_modern(axis.title.space = 10) +
  theme(strip.text = element_text(size = 10),
        strip.background = element_rect(fill = c("grey"), color = "white"))

p3

model <- brms::brm(Confidence ~ Answer / Condition / Intero_Focus + (1|Participant) + (1|Item), 
                   data = df, family = "beta", seed=3, refresh=0,
                   prior = set_prior("student_t(1, 0, 1)", class = "b"))
link_data <- estimate_relation(model, at = c("Condition", "Answer", "Intero_Focus"), length = 30)

p4 <- ggplot(link_data, aes(x = Intero_Focus, y = Predicted)) +
  geom_segment(data = data,
                  aes(x = Intero_Focus_Mean,
                      xend = Intero_Focus_Mean,
                      y = Confidence_low,
                      yend = Confidence_high,
                      color = Condition),
                  alpha=1/3) +
  geom_point2(data = data,
                  aes(x = Intero_Focus_Mean,
                      y = Confidence_Mean,
                      color = Condition),
              size=2) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Condition), alpha = 1/3) +
  geom_line(aes(color = Condition), size=1) +
  geom_text(data=data.frame(Intero_Focus = -0.2,
                            Predicted = c(0.55, 0.75),
                            Condition = c("Polygraph", "Interrogation"),
                            Answer = c("Truth", "Truth")),
            label = c("*", "*"),
            color = c("#2196F3", "#FF5722"),
            size = 10, show.legend = FALSE, fontface = "bold") +
  labs(y = "Confidence",
       x = expression(Interoception["  (.87 MAIA Not-distracting, -.40 MAIA Emotional Awareness, .33 HCT Accuracy)"])) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  scale_fill_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  facet_wrap(~Answer) +
  theme_modern(axis.title.space = 10) +
  theme(strip.text = element_text(size = 10),
        strip.background = element_rect(fill = c("grey"), color = "white"))
p4

model <- brms::brm(Confidence ~ Answer / Condition / Intero_Regulation + (1|Participant) + (1|Item), 
                   data = df, family = "beta", seed=3, refresh=0,
                   prior = set_prior("student_t(1, 0, 1)", class = "b"))

link_data <- estimate_relation(model, at = c("Condition", "Answer", "Intero_Regulation"), length = 30)

p5 <- ggplot(link_data, aes(x = Intero_Regulation, y = Predicted)) +
  geom_segment(data = data,
                  aes(x = Intero_Regulation_Mean,
                      xend = Intero_Regulation_Mean,
                      y = Confidence_low,
                      yend = Confidence_high,
                      color = Condition),
                  alpha=1/3) +
  geom_point2(data = data,
                  aes(x = Intero_Regulation_Mean,
                      y = Confidence_Mean,
                      color = Condition),
              size=2) +
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Condition), alpha = 1/3) +
  geom_line(aes(color = Condition), size=1) +
  geom_text(data=data.frame(Intero_Regulation = -0.15, 
                            Predicted = c(0.25, 0.45, 0.75),
                            Condition = c("Polygraph", "Interrogation", "Polygraph"), 
                            Answer = c("Lie", "Lie", "Truth")),
            label = c("***", "*", "*"),
            color = c("#FF5722", "#2196F3", "#FF5722"),
            size = 10, show.legend = FALSE, fontface = "bold") +
  labs(y = "Confidence",
       x = expression(Interoception["  (.71 MAIA Not-worrying, .61 HCT Accuracy, .40 MAIA Trusting)"])) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  scale_fill_manual(values = c("Polygraph" = "#FF5722", "Interrogation" = "#2196F3")) +
  facet_wrap(~Answer) +
  theme_modern(axis.title.space = 10) +
  theme(strip.text = element_text(size = 10),
        strip.background = element_rect(fill = c("grey"), color = "white")) 
p5

p <- (p1 / p2 / p3 / p4 / p5) + plot_layout(guides = "collect") + plot_annotation(title = "Interindividual correlates of the confidence\nthat one tells a convincing lie", theme = list(plot.title = element_text(face = "bold", hjust = 0.5)))
p
ggsave("figures/Figure1.png", width=12, height=15)
```

### 2022-10-04 paper/manuscript.Rmd
update

---
title: '**The Role of Interoception and Theory of Mind in Deception**'
shorttitle        : "TITLE"
author:
  - name          : "Dominique Makowski"
    affiliation   : " 1, *"
    address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
    email         : "dmakowski@ntu.edu.sg"
  - name          : "Zen J. Lau"
    affiliation   : " 1"
  - name          : "Tam Pham"
    affiliation   : " 1"
  - name          : "S.H. Annabel Chen"
    affiliation   : " 1, 2, 3, 4"
    corresponding : yes    # Define only one corresponding author
    address       : "HSS 04-19, 48 Nanyang Avenue, Singapore"
    email         : "annabelchen@ntu.edu.sg"
affiliation:
  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
  - id            : "4"
    institution   : "National Institute of Education, Nanyang Technological University, Singapore"
abstract: |
output:
  papaja::apa6_word:
    keep_tex: yes
  pdf_document: default
  papaja::apa6_pdf:
    keep_tex: yes
keywords: ""
wordcount: ''
bibliography: ["references.bib"]
floatsintext: yes
figsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
header-includes:
- \usepackage[labelfont=bf, font={color=gray,small}]{caption}
- \usepackage{float}
- \usepackage[document]{ragged2e}
csl: utils/apa.csl
---

@@

The data analysis was carried out using R.... [CITE STUFF]. The full reproducible analysis script is available [INSERT LINK].

We implimented a linear mixed-effect model where

From the deception task, there were three dependent variables: confidence, reaction time and heart rate. For each dependent variable, we investigated its relationship with theory of mind and interoception abilities as measured by the tests on the previous slide. Participant and the specific question asked were added in as random effects, as they are not independent factors. For example, certain participants may be more confident in lying unrelated to their theory of mind or interoception abilities, or certain questions may have a longer reaction time

To further build on the model, trials for lies and truths were separated using an embedded model and the interaction between the dependent variable and condition was probed. This reveals instances where the effect is greater in one condition, for example a higher total Yoni Task score is correlated with decreased lie confidence, but only in the polygraph condition.

### Data Pre-Processing and Analyses

@@

INSERT VERY FANCY, VERY PRETTY GRAPHS

Implicit relationship between dependent variables. Inter-Measure Correlation:

-   (confident and reaction time) When lying, the faster they answer, the more confident they are.

-   less confident in lies than truths (expected) Condition interaction, greater effect in polygraph (relying on interoception) likely less intuitive feedback cues?? Slower in polygraph makes sense if less intuitive

-   (HR and RT) no relationship

TOM correlated with trials in polygraph (interoception), not expected(!) because expected it to correlate with Interrogation

higher HCT accuracy and MAIA positively correlated with lie confidence in polygraph condition

No effect for reaction time

BES positively correlated with HR,

HCT accuracy negatively correlated with HR

>>>>

---
title: '**The Heart Lies: The Role of Interoception and Theory of Mind in Deception**'
shorttitle        : "TITLE"
author:
  - name          : "Dominique Makowski"
    affiliation   : " 1, *"
    address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
    email         : "dmakowski@ntu.edu.sg"
  - name          : "Zen J. Lau"
    affiliation   : " 1"
  - name          : "Tam Pham"
    affiliation   : " 1"
  - name          : "S.H. Annabel Chen"
    affiliation   : " 1, 2, 3, 4"
    corresponding : yes    # Define only one corresponding author
    address       : "HSS 04-19, 48 Nanyang Avenue, Singapore"
    email         : "annabelchen@ntu.edu.sg"
affiliation:
  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
  - id            : "4"
    institution   : "National Institute of Education, Nanyang Technological University, Singapore"
abstract: |
output:
  papaja::apa6_word:
    keep_tex: yes
  pdf_document: default
  papaja::apa6_pdf:
    keep_tex: yes
keywords: ""
wordcount: ''
bibliography: ["references.bib"]
floatsintext: yes
figsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
header-includes:
- \usepackage[labelfont=bf, font={color=gray,small}]{caption}
- \usepackage{float}
- \usepackage[document]{ragged2e}
csl: utils/apa.csl
---

@@

The manipulation checks consist of testing whether there is an effect of question phrasing (direct vs. indirect), and condition (polygraph *vs.* interrogation), on our 3 outcome variables: the participants' ratings of confidence that their answers (lies *vs.* truths) were convincing, the response time (RT), and the heart rate change associated with the response. This analysis was performed using mixed models with the participants and questions both entered as random factors. Marginal contrasts analysis was also performed to clarify the differences between conditions. To allow for a better quantification of the uncertainty associated with the effects, as well increase the robustness to outliers and artifactual findings, all statistics were undertaken under the Bayesian framework [**ref makowski2019existence**], using informative priors centred around 0 ($t_{Confidence}(1, 0, 1)$, $t_{RT}(1, 0, 3)$, $t_{Heartrate}(1, 0, 8)$).

To further counterbalance the low number of participants and maximize the signal-to-noise ratio, we performed a feature reduction on our two groups of predictor variables (namely, theory of mind and interoception). We used factor analysis (using the method agreement procedure to estimate the optimal number of dimensions, **REF PARAMETERS PKG**), rather PCA, as the goal was to extract meaningful and consistent factors, rather than merely maximizing the variance explained. Then, we modelled the relationship between these inter-individual composite scores (the analysis for all the variables is included in Supplementary Materials) and the 3 outcome variables in interaction with the condition (polygraph *vs.* interrogation). Finally, we also investigated the relationship between the deception scale traits, and the theory of mind and interoception scores using Bayesian correlations.

The data analysis was carried out using R, the *brms* package [*REF*] and the *easystats* ecosystem **[REFS]**. As the full reproducible analysis script and statistical results are available at **[INSERT LINK]**, we will only focus on significant results in the manuscript.

@@

## Manipulation Check

Compared to truth, lies were rated with less confidence ($\beta=-1.09,~95\%~CI~[-0.94, -1.25],~pd = 100\%$). Moreover, lies told in the polygraph condition were rated with less confidence ($\delta=-0.24,~95\%~CI~[-0.10, -0.39],~pd > 99\%$).
The RT did not differ between truth and lies, but was significantly slower in the polygraph condition ($\beta=0.14,~95\%~CI~[0.03, 0.25],~pd > 99\%$) for both conditions.
The heart rate was significantly more elevated during lies as compared to truth ($\beta=1.31,~95\%~CI~[2.14, 0.50],~pd > 99.9\%$), and during interrogation as compared to the polygraph condition ($\beta=4.99	,~95\%~CI~[5.82, 4.19],~pd = 100\%$).

The indirect phrasing of the question only had a significant effect on RT ($\beta=0.34,~95\%~CI~[0.23, 0.44],~pd = 100\%$), leading to slower answers, regardless of whether they were lies or truths.

## Feature Reduction

The 3 YONI-task dimensions and the 2 BES traits were combined into a unique factor, labelled *Theory of Mind* (explaining 35.76\% of variance). It was loaded by the cognitive (.89), affective (.77), physical (.45) YONI dimension, and the affective (.41) and cognitive (.17) traits of the BES.

The 8 MAIA dimensions and the 3 HCT components were reduced to 4 factors (explaining 65.17\% of variance). The first factor, labelled *Interoception - Meta* (23.59\%), was loaded primarily by Attention Regulation (.97), Self-regulation (.63), Emotional awareness (.60), and noticing (.49) dimensions of the MAIA and the HCT confidence score (.40). The second factor, labelled *Interoception - Listening* (18.54\%), was primarily loaded by the body listening (.92) and trusting (.53) MAIA dimensions, and the awareness (-.60) and confidence (.46) HCT scores. The third factor, labelled *Interoception - Focus* (12.07\%), was primarily loaded by MAIA not-distracting (.87), emotional awareness (-.40) and HCT accuracy (.33). The fourth factor, labelled *Interoception - Regulation* (10.97\%), was primarily loaded by MAIA not-worrying (.71), HCT accuracy (.61) and MAIA trusting (.40). However, due to the interpretation difficulty, we will refer to these dimensions with their main correlates in **Figure 1**.

## Theory of Mind

The composite *ToM* score was significantly associated with less confident ($\beta=-0.19,~95\%~CI~[-0.36, -0.02],~pd = 98.47\%$) and slower lies ($\beta=0.42,~95\%~CI~[0.01, 0.83],~pd = 97.67%\%$), specifically in the polygraph condition. No significant effect was found heart rate, and no correlation with the deception scale traits.

## Interoception

The *Meta* interoception score was significantly associated with an increased confidence in lies, specifically in the polygraph condition ($\beta=0.20,~95\%~CI~[0.03, 0.35],~pd = 98.98\%$). It was also associated with faster answers for both lies ($\beta=-0.54,~95\%~CI~[-0.93, -0.15],~pd = 99.67\%$) and truths ($\beta=-0.29,~95\%~CI~[-0.63, 0.03],~pd = 96.10%\%$), specifically in the polygraph condition. No significant association was found with heart rate.

The *Listening* interoception score was significantly associated with an increased confidence in lies, in the polygraph ($\beta=0.43,~95\%~CI~[0.27, 0.59],~pd = 100\%$) and interrogation condition ($\beta=0.16,~95\%~CI~[0.01, 0.32],~pd = 98.04\%$). It was also associated with faster answers, particilarly for lies ($\beta=-0.42,~95\%~CI~[-0.82, -0.03],~pd = 98.19\%$) and truths ($\beta=-0.36,~95\%~CI~[-0.76, 0.03],~pd = 96.49\%$) in the polygraph condition. No significant association was found with heart rate. This score also correlated with the deception *Contextuality* trait ($r = 0.50,~95\%~CI~[0.04, 0.64],~BF_{10} = 3.48\%$).

The *Focus* interoception score was significantly associated with an increased confidence in truthful responses in the polygraph ($\beta=0.17,~95\%~CI~[-0.01, 0.34],~pd = 97.16\%$) and interrogation condition ($\beta=0.15,~95\%~CI~[-0.02, 0.32],~pd = 95.76%\%$). No significant association was found with RT or heart rate, but the score correlated with the deception *Ability* trait ($r = 0.50,~95\%~CI~[0.22, 0.74],~BF_{10} = 34.37\%$).

The *Regulation* interoception score was significantly associated with an increased confidence in lies in the polygraph ($\beta=0.32,~95\%~CI~[0.14, 0.51],~pd = 99.99\%$) and the interrogation conditions ($\beta=0.18,~95\%~CI~[0.00, 0.36],~pd = 97.42\%$), and with a decreased confidence in truth only in the polygraph condition ($\beta=-0.1,~95\%~CI~[-0.36, 0.01],~pd = 97.16\%$). No significant association was found with RT or heart rate.
